{ \fontfamily{ptm}\selectfont First Order Nonsmooth Optimization: Algorithm Design, Variational analysis, and Applications }

### 2025-01-08 Notebook LaTEX/Proposal/main.tex
>w<, changes, sync

\section{Introduction}
    Let $\RR^n$ be the ambient space. We consider 
    \begin{align}
        \min_{x \in \RR^n} \left\lbrace
            F(x): f(x) + g(x)
        \right\rbrace.
    \end{align}\label{eqn:additive-comp-obj}
    Unless specified, assume $f:\RR^n \rightarrow \RR$ is $L$-Lipscthiz smooth $\mu \ge 0$ strongly convex and $g:Q \rightarrow \overline \RR$ is convex. 
    This type of problem is referred to as additive composite problems in the literature. 
    \par
    Our ongoing research concerns accelerated proximal gradient type method for solving (\ref{eqn:additive-comp-obj}). 
    In the expository writing by Walkington \cite{noel_nesterovs_nodate}, a variant for of accelerated gradient method for strongly convex function $f$ is discussed. 
    We had some lingering questions after reading it. 
    \begin{enumerate}
        \item Do there exist a unified description for the convergence for both variants of the algorithms?
        \item Is it possible to attain fast convergence rate without knowledge about the strong convexity of function $f$?
        \item Is it possible to describe the convergence of function value for momentum sequences that are much weaker than the Nesterov's rule? 
    \end{enumerate}
    The good news is we have definitive answers for all questions by our own efforts of research. 
    Section \ref{sec:unify-nes-acceleration}, \ref{sec:free-rwapg} are our ongoing research which present the answers to the questions. 
    \par
    In Section \ref{sec:unify-nes-acceleration}, we proposed the method of ``Relaxed Weak Accelerated Proximal Gradient (R-WAPG)'' as the foundation to describe several variants of Accelerated proximal gradient method in the literatures. 
    The convergence theories of R-WAPG allows us to model convergence of accelerated proximal gradient method where the momentum sequence doesn't strictly follow the conditions presented in the literatures. 
    The descriptive power of R-WAPG allows convergence analysis for all the variants using one single theorem. 
    \par
    In Section \ref{sec:free-rwapg} we propose a practical algorithm that exploits a specific term in the proof of R-WAPG to achieve faster convergence for solving (\ref{eqn:additive-comp-obj}) without knowing parameter $L, \mu$ in prior. 
    Results of numerical experiments are presented. 
    \par
    Section \ref{sec:catalyst} are results of literatures review in MATH 590. 
    \todo{
        $\blacksquare$ Add Lin's papers and Paquette's papers.
    }
    {\hl{It's based on a series of papers}} in the topic of Catalyst Meta Acceleration method for First Order Variance Reduced Methods. 
    We will point out potential future direction of research of Catalyst acceleration. 
    \par
    Section \ref{sec:inexact-prox}, \ref{sec:nes-acc-ncnvx} preview literatures in nonsmooth optimization frontier research where progress and impacts can be made.  
    \subsection{Theme of the research}
        This section specifies a theme of the research in this proposal. 
        Out first objective is to explore the Goldilocks zones between these topics: theories of variational analysis, design of continuous optimization algorithm and applications in sciences, engineering, and statistics.  
        Our second objective is to identify the ``chemistry'' occurring between properties of functions and the designs of continuous optimizations algorithm and how it impacts the convergence and behaviors of the algorithms.

@@

        {\hl{incremental methods}}

@@

        In each iteration, only the gradient of a few samples are accessed and uses simple calculations from previously stored gradient to attain overall better complexity 
        It can achieve that because the variance of the estimated gradients are smaller when the iterates are close to the set of true minimizer. 

@@

                \item 

>>>>

\section{Introduction}
    Let $\RR^n$ be the ambient space. We consider 
    \begin{align}
        \min_{x \in \RR^n} \left\lbrace
            F(x): f(x) + g(x)
        \right\rbrace.
    \end{align}\label{eqn:additive-comp-obj}
    Unless specified, assume $f:\RR^n \rightarrow \RR$ is $L$-Lipscthiz smooth $\mu \ge 0$ strongly convex and $g:Q \rightarrow \overline \RR$ is convex. 
    This type of problem is referred to as additive composite problems in the literature. 
    \par
    Our ongoing research concerns accelerated proximal gradient type method for solving (\ref{eqn:additive-comp-obj}). 
    In the expository writing by Walkington \cite{noel_nesterovs_nodate}, a variant for of accelerated gradient method for strongly convex function $f$ is discussed. 
    We had some lingering questions after reading it. 
    \begin{enumerate}
        \item Do there exist a unified description for the convergence for both variants of the algorithms?
        \item Is it possible to attain fast convergence rate without knowledge about the strong convexity of function $f$?
        \item Is it possible to describe the convergence of function value for momentum sequences that are much weaker than the Nesterov's rule? 
    \end{enumerate}
    The good news is we have definitive answers for all questions by our own efforts of research. 
    \todoinline{\noindent
        $\square$ Change all the organization parts here
    }
    Section \ref{sec:unify-nes-acceleration}, \ref{sec:free-rwapg} are our ongoing research which present the answers to the questions. 
    \par
    In Section \ref{sec:unify-nes-acceleration}, we proposed the method of ``Relaxed Weak Accelerated Proximal Gradient (R-WAPG)'' as the foundation to describe several variants of Accelerated proximal gradient method in the literatures. 
    The convergence theories of R-WAPG allows us to model convergence of accelerated proximal gradient method where the momentum sequence doesn't strictly follow the conditions presented in the literatures. 
    The descriptive power of R-WAPG allows convergence analysis for all the variants using one single theorem. 
    \par
    In Section \ref{sec:free-rwapg} we propose a practical algorithm that exploits a specific term in the proof of R-WAPG to achieve faster convergence for solving (\ref{eqn:additive-comp-obj}) without knowing parameter $L, \mu$ in prior. 
    Results of numerical experiments are presented. 
    \par
    Section \ref{sec:catalyst} are results of literatures review in MATH 590. 
    \todo{
        $\blacksquare$ Add Lin's papers and Paquette's papers.
    }
    {\hl{It's based on a series of papers}} in the topic of Catalyst Meta Acceleration method for First Order Variance Reduced Methods. 
    We will point out potential future direction of research of Catalyst acceleration. 
    \par
    Section \ref{sec:inexact-prox}, \ref{sec:nes-acc-ncnvx} preview literatures in nonsmooth optimization frontier research where progress and impacts can be made.  
    \subsection{Theme of the research}
        This section specifies a theme of the research in this proposal. 
        Out first objective is to explore the Goldilocks zones between these topics: theories of variational analysis, design of continuous optimization algorithm and applications in sciences, engineering, and statistics.  
        Our second objective is to identify the ``chemistry'' occurring between properties of functions and the designs of continuous optimizations algorithm and how it impacts the convergence and behaviors of the algorithms.

@@

        {\hl{incremental methods}} \cite{bertsekas_incremental_2011,bertsekas_incremental_2017}

@@

        In each iteration, only the gradient of a few samples are accessed and uses simple calculations from previously stored gradient to attain overall better complexity.
        Compare to traditional stochastic gradient, a smaller variance of the estimated gradients near the minimizer gives faster convergence rate. 
        Major examples of VRMs include SVRG by Xiao, Zhang \cite{xiao_proximal_2014}, Finito by Defazio et al. \cite{defazio_finito_2014}, SAG by Schmidt et al. \cite{schmidt_minimizing_2017}, and SAGA by \cite{defazio_saga_2014}. 

@@

                \item[$\blacksquare$] SVRG: Proximal stochastic gradient with progressive variance reduction. 
                \item[$\blacksquare$] Finito: A faster, permutable incremental gradient method for big data problems
                \item[$\blacksquare$] SAG: Minimizing finite sums with the stochastic average gradient
                \item[$\blacksquare$] SAGA: A fast incremental gradient method with support for non-strongly convex composite objectives. 
